Longitudinal models

### 2024-12-13 5_longitudinal.qmd
paper and longitudinal updates

Notes from Nilam:
Look at this part of the latent growth tutorial #observed residual variances (constrained to equality)- math2 ~~ sigma2_u*math2

the two-wave model:
change in inter-individual differences
vs. differences in intra-individual change

@@

```{r}
library(here)
source(here("helper","common.R"))
library(lavaan)
library(tidySEM)

@@

# Reliability

How to do this right? WRONG RIGHT NOW

```{r}
MONTH_CUTOFF <- 3

d_reliability <- d_sub_long |>
  filter(admin_num <= 2) |>
  pivot_wider(id_cols = c("subject_id","dataset_name"), 
              names_from = "admin_num",
              values_from = c("log_rt", "log_rt_var", "long_window_accuracy", 
                              "long_window_acc_var", "prod", "comp"))

dataset_reliabilities <- d_reliability |>
  group_by(dataset_name) |>
  summarise(rt = cor(log_rt_1, log_rt_2, use = "pairwise.complete.obs"),
            acc = cor(long_window_accuracy_1, 
                          long_window_accuracy_2, use = "pairwise.complete.obs")) |>
  filter(dataset_name != "fernald_totlot") 
```

# Longitudinal factor analysis

```{r}
d_sub_s <- d_sub |>
  ungroup() |>
  select(dataset_name, subject_id, administration_id, age, 
         time_since_t0, delta_t,
         log_rt, log_rt_var, 
         long_window_accuracy, long_window_acc_var, prod, comp, ) |>
  rename(acc = long_window_accuracy, 
         acc_sd = long_window_acc_var, 
         log_rt_sd = log_rt_var) |>
  ungroup() |>
  mutate(across(all_of(c("log_rt", "log_rt_sd", "acc", "acc_sd", 
                         "prod", "comp")), 
                       ~ age_scale(.x, age))) 
```

@@

# Two-time point

@@

```
```{r}
layout_long = matrix(nrow=7, ncol = 7,
                     data = c("log_rt_sd_t1","log_rt_t1", "acc_t1", "acc_sd_t1", "prod_t1","comp_t1", NA,
                               NA, NA, NA, NA, NA, NA,  NA,
                              "speed_t1",NA,"accuracy_t1",NA,"vocab_t1",NA,  NA,
                              NA, NA, NA, NA, NA, NA,  NA,
                               NA,"speed_t2",NA,"accuracy_t2",NA,NA,"vocab_t2",
                               NA, NA, NA, NA, NA, NA,  NA,
                               NA,"log_rt_sd_t2","log_rt_t2", "acc_t2", "acc_sd_t2", "prod_t2", "comp_t2"), 
                     byrow = TRUE)
graph_sem(model = fit3_long, text_size = 2, layout = t(layout_long))
```

# Full longitudinal model

@@

d_sub_wide |>
  ungroup() |>
  summarise(across(log_rt_t1:comp_t4, 
                   ~sum(!is.na(.x))))
```

@@

vocab_t3 ~~ accuracy_t3
vocab_t3 ~~ speed_t3
accuracy_t3 ~~ speed_t3

vocab_t4 ~~ accuracy_t4
vocab_t4 ~~ speed_t4
accuracy_t4 ~~ speed_t4

# means for the latents
vocab_t1 ~ 1
vocab_t2 ~ 1
vocab_t3 ~ 1
vocab_t4 ~ 1
accuracy_t1 ~ 1
accuracy_t2 ~ 1
accuracy_t3 ~ 1
accuracy_t4 ~ 1
speed_t1 ~ 1
speed_t2 ~ 1
speed_t3 ~ 1
speed_t4 ~ 1

@@

# # residual variance for the latents
# vocab_t1 ~~ NA*vocab_t1
# vocab_t2 ~~ NA*vocab_t2
# accuracy_t1 ~~ NA*accuracy_t1
# accuracy_t2 ~~ NA*accuracy_t2
# speed_t1 ~~ NA*speed_t1
# speed_t2 ~~ NA*speed_t2
"

@@

```{r}
layout_dt = matrix(nrow=11, ncol = 7,
                   data = c("prod_t1", "prod_t2", "prod_t3", "prod_t4", "vocab_intercept", NA, NA,
                            NA, NA, NA, NA, NA, NA, NA,
                            "comp_t1", NA, NA, NA, "vocab_slope", NA,NA,
                            NA, NA, NA, NA, NA, NA, NA,
                            "accuracy_t1", "accuracy_t2", "accuracy_t3", "accuracy_t4", "accuracy_intercept", NA, NA,
                            NA, NA, NA, NA, NA, NA, NA,
                            "acc_sd_t1", "acc_sd_t2", "acc_sd_t3", "accuracy_sd_t4", "accuracy_slope", NA, NA,
                            NA, NA, NA, NA, NA, NA, NA,
                            "log_rt_t1", "log_rt_t2", "log_rt_t3", "log_rt_t4", "speed_int", NA, NA,
                            NA, NA, NA, NA, NA, NA, NA,
                            "log_rt_sd_t1", "log_rt_sd_t2", "log_rt_sd_t3", "log_rt_sd_t4", "speed_slope", NA, NA),
                   byrow = TRUE)

graph_sem(model = fit3_long, text_size = 3, layout = layout_dt)

@@

>>>>

@@

```{r}
library(here)
source(here("helper","common.R"))
library(lavaan)
library(tidySEM)
library(nlme)
library(emmeans)
library(lme4)
library(lmerTest)

@@

# Draft 1: Replicate growth models from Fernald & Marchman 2012

First, we try replicating the original growth models.

## Just FM2012

```{r}
fm2012 <- filter(d_sub, dataset_name == "fernald_marchman_2012") |>
  group_by(subject_id) |>
  mutate(age_group = case_when(age < 21 ~ 18, 
                               age > 21 & age < 27 ~ 24, 
                               .default = 30)) |>
  arrange(age) |>
  mutate(rt_t0 = rt[1]) |>
  ungroup() |>
  mutate(rt_t0_group = ifelse(rt_t0 < median(rt_t0), "low", "high"))
```

```{r}
ggplot(fm2012, 
       aes(x = age_group, y= prod)) + 
  geom_line(alpha = .2, aes(group = subject_id)) + 
  stat_summary(aes(col = rt_t0_group)) + 
  geom_smooth(aes(col = rt_t0_group), method = "lm", formula = y ~ poly(x, 2))
```

This looks similar (but maybe a bit worse with log rt.

```{r}
fm_mod <- lmer(prod ~ age * rt_t0 + (age | subject_id), 
     data = fm2012, 
     control = lmerControl(optimizer = "bobyqa"))

summary(fm_mod)
```

Note that this model looks a lot better with quadratic growth. (They used quadratics in the original).

```{r}
fm_mod <- lmer(prod ~ age * rt_t0 + I(age^2) * rt_t0 + (age | subject_id), 
     data = fm2012, 
     control = lmerControl(optimizer = "bobyqa"))

summary(fm_mod)
```

## All data

Let's try the same model with all the data.

```{r}
d_growth <- d_sub |>
  filter(!is.na(prod), subject_id %in% longitudinal$subject_id) |>
  group_by(subject_id) |>
  arrange(age) |>
  mutate(rt_t0 = log_rt[1]) |>
  ungroup() |>
  mutate(rt_t0_group = ifelse(rt_t0 < median(rt_t0, na.rm=TRUE), "low", "high"))

ggplot(filter(d_growth, !is.na(rt_t0)), 
              aes(x = age, y= prod)) +
  geom_line(alpha = .2, aes(group = subject_id)) +
  geom_smooth(aes(col = rt_t0_group), method = "lm", formula = y ~ x + I(x^2))
```

```{r}
ggplot(filter(d_growth, !is.na(rt_t0)), 
              aes(x = age, y= prod)) +
  geom_line(alpha = .2, aes(group = subject_id)) +
  geom_smooth(aes(col = rt_t0_group), method = "lm", formula = y ~ x + I(x^2))
```

Linear.

```{r}
d_growth$age_15 <- d_growth$age - 15
all_mod <- lmer(prod ~ age_15 *rt_t0 +  (age  | subject_id) + (age | dataset_name), 
     data = d_growth, 
     control = lmerControl(optimizer = "bobyqa"))

summary(all_mod)
```

Again better with quadratics. (Won't converge with nested quadratics in random effects).

```{r}
d_growth$age_15 <- d_growth$age - 15
all_mod <- lmer(prod ~ age_15 *rt_t0  + I(age_15^2) * rt_t0 + (age  | subject_id) + (age | dataset_name), 
     data = d_growth, 
     control = lmerControl(optimizer = "bobyqa"))

summary(all_mod)
```

So what we see here is that RT at t0 is very predictive of something about growth.

On the other hand, these models don't take into account RT at t1 or t2. So they don't tell us if RT is a stable trait that's associated with growth...

# Draft 2: Two-time points

So we try fitting two factor analyses and looking at coupling over time.

```{r}
d_sub_s <- d_sub |>
  ungroup() |>
  select(dataset_name, subject_id, administration_id, age, 
         time_since_t0, delta_t, admin_num,
         log_rt, log_rt_var, 
         long_window_accuracy, long_window_acc_var, prod, comp, ) |>
  rename(acc = long_window_accuracy, 
         acc_sd = long_window_acc_var, 
         log_rt_sd = log_rt_var) |>
  ungroup() |>
  mutate(across(all_of(c("log_rt", "log_rt_sd", "acc", "acc_sd", 
                         "prod", "comp")), 
                       ~ age_scale(.x, age))) 
```

@@

@@

summary(fit3_long, fit.measures=TRUE, standardize=TRUE)
```
```{r}
layout_long = matrix(nrow=7, ncol = 7,
                     data = c("log_rt_sd_t1","log_rt_t1", "acc_t1", "acc_sd_t1", "prod_t1","comp_t1", NA,
                               NA, NA, NA, NA, NA, NA,  NA,
                              "speed_t1",NA,"accuracy_t1",NA,"vocab_t1",NA,  NA,
                              NA, NA, NA, NA, NA, NA,  NA,
                               NA,"speed_t2",NA,"accuracy_t2",NA,NA,"vocab_t2",
                               NA, NA, NA, NA, NA, NA,  NA,
                               NA,"log_rt_sd_t2","log_rt_t2", "acc_t2", "acc_sd_t2", "prod_t2", "comp_t2"), 
                     byrow = TRUE)
graph_sem(model = fit3_long, text_size = 2, layout = t(layout_long))
```

# Draft 3: Full longitudinal model

@@

d_sub_wide |>
  ungroup() |>
  summarise(across(log_rt_t1:comp_t4, 
                   ~sum(!is.na(.x))))
```
## with just the observed variables

```{r}
slopes_model_long <- "

# regressions
accuracy_intercept =~ 1*acc_t1 + 1*acc_t2 + 1*acc_t3 + 1*acc_t4 
accuracy_slope =~ 1*acc_t1 + 2*acc_t2 + 3*acc_t3 + 4*acc_t4
speed_intercept =~ 1*log_rt_t1 + 1*log_rt_t2 + 1*log_rt_t3 + 1*log_rt_t4 
speed_slope =~ 1*log_rt_t1 + 2*log_rt_t2 + 3*log_rt_t3 + 4*log_rt_t4
vocab_intercept =~ 1*prod_t1 + 1*prod_t2 + 1*prod_t3 + 1*prod_t4 
vocab_slope =~ 1*prod_t1 + 2*prod_t2 + 3*prod_t3 + 4*prod_t4
"

slopes_long <- growth(slopes_model_long, d_sub_wide, std.lv=TRUE, missing='fiml')

summary(slopes_long, fit.measures=TRUE, standardize=TRUE)

```{r}
layout <- read.csv(here("misc/layout_slopes.csv"), header = FALSE)

graph_sem(model = slopes_long, text_size = 3, layout=t(layout))

This is still linear. We can add quadratic?

```{r}
d_sub_wide |>
  select(subject_id, prod_t1, prod_t2, prod_t3, prod_t4) |>
  pivot_longer(starts_with("prod"), names_to = "variable", values_to = "prod") |>
  mutate(t = as.numeric(str_sub(variable, -1))) |>
  ggplot(aes(x = t, y = prod)) + 
  geom_line(aes(group = subject_id), alpha = .1) + 
  geom_smooth(aes(group = 1), method = "lm") + 
  geom_smooth(aes(group = 1), method = "lm", formula = y ~ poly(x, 2), col = "green")
```

But it seems unnecessary for this view of the data. Arguably we're losing a lot by doing it this way.

## try finer-grained rebinning

Let's try putting prod back to its natural units?

```{r}
d_sub_wide_fine <- d_sub_s |>
  mutate(t = cut(time_since_t0, breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 30), 
                 include.lowest = TRUE)) |>
  filter(t != "(12,30]") |>
  mutate(t = as.numeric(t)) |>
  group_by(subject_id, dataset_name, t) |>
  summarise(across(all_of(c("log_rt", "log_rt_sd", "acc",
                            "acc_sd", "prod", "comp")),
                   ~ mean(.x, na.rm=TRUE))) |> 
  mutate(across(all_of(c("log_rt", "log_rt_sd", "acc",
                         "acc_sd", "prod", "comp")),
                ~ ifelse(is.nan(.x), NA, .x))) |>
  pivot_wider(id_cols = c("subject_id","dataset_name"), 
              names_from = "t",
              values_from = c("log_rt", "log_rt_sd", "acc", 
                              "acc_sd", "prod", "comp"), 
              names_prefix = "t")

d_sub_wide_fine |>
  ungroup() |>
  summarise(across(log_rt_t1:comp_t4, 
                   ~sum(!is.na(.x))))
```

```{r}
d_sub_wide_fine |>
  select(subject_id, starts_with("prod")) |>
  pivot_longer(starts_with("prod"), names_to = "variable", values_to = "prod") |>
  mutate(t = as.numeric(str_sub(variable, -1))) |>
  ggplot(aes(x = t, y = prod)) + 
  geom_line(aes(group = subject_id), alpha = .1) 
```

```{r}
slopes_model_long <- "

# regressions
accuracy_intercept =~ 1*acc_t1 + 1*acc_t2 + 1*acc_t3 + 1*acc_t4 + 1*acc_t5 
accuracy_slope =~ 1*acc_t1 + 2*acc_t2 + 3*acc_t3 + 4*acc_t4 + 5*acc_t5 
speed_intercept =~ 1*log_rt_t1 + 1*log_rt_t2 + 1*log_rt_t3 + 1*log_rt_t4 + 1*log_rt_t5 
speed_slope =~ 1*log_rt_t1 + 2*log_rt_t2 + 3*log_rt_t3 + 4*log_rt_t4 + 5*log_rt_t5
vocab_intercept =~ 1*prod_t1 + 1*prod_t2 + 1*prod_t3 + 1*prod_t4 + 1*prod_t5 
vocab_slope =~ 1*prod_t1 + 2*prod_t2 + 3*prod_t3 + 4*prod_t4 + 5*prod_t5

# let variances of each latent vary
accuracy_intercept ~~ NA*accuracy_intercept
accuracy_slope ~~ NA*accuracy_slope
speed_intercept ~~ NA*speed_intercept
speed_slope ~~ NA*speed_slope
vocab_intercept ~~ NA*vocab_intercept
vocab_slope ~~ NA*vocab_slope
"

slopes_long <- growth(slopes_model_long, d_sub_wide_fine, std.lv=TRUE, missing='fiml')

summary(slopes_long, fit.measures=TRUE, standardize=TRUE)

```{r}
layout <- read.csv(here("misc/layout_slopes_5.csv"), header = FALSE)

graph_sem(model = slopes_long, text_size = 3, layout=t(layout))

## with lots of latent factor structure

This one still seems not quite right.

@@

# means for the latents
# vocab_t1 ~ 1
# vocab_t2 ~ 1
# vocab_t3 ~ 1
# vocab_t4 ~ 1
# accuracy_t1 ~ 1
# accuracy_t2 ~ 1
# accuracy_t3 ~ 1
# accuracy_t4 ~ 1
# speed_t1 ~ 1
# speed_t2 ~ 1
# speed_t3 ~ 1
# speed_t4 ~ 1

# autoregression for the latents
# vocab_t2 ~ 1 * vocab_t1
# vocab_t3 ~ 1 * vocab_t2
# vocab_t4 ~ 1 * vocab_t3
# accuracy_t2 ~ 1 * accuracy_t1
# accuracy_t3 ~ 1 * accuracy_t2
# accuracy_t4 ~ 1 * accuracy_t3
# speed_t2 ~ 1 * speed_t1
# speed_t3 ~ 1 * speed_t2
# speed_t4 ~ 1 * speed_t3

@@

# residual variance for the latents
vocab_t1 ~~ NA*vocab_t1
vocab_t2 ~~ NA*vocab_t2
vocab_t3 ~~ NA*vocab_t3
vocab_t4 ~~ NA*vocab_t4
accuracy_t1 ~~ NA*accuracy_t1
accuracy_t2 ~~ NA*accuracy_t2
accuracy_t3 ~~ NA*accuracy_t3
accuracy_t4 ~~ NA*accuracy_t4
speed_t1 ~~ NA*speed_t1
speed_t2 ~~ NA*speed_t2
speed_t3 ~~ NA*speed_t3
speed_t4 ~~ NA*speed_t4

# residual variance for the other latents
vocab_intercept ~~ NA*vocab_intercept
vocab_slope ~~ NA*vocab_slope
accuracy_intercept ~~ NA*accuracy_intercept
accuracy_slope ~~ NA*accuracy_slope
speed_intercept ~~ NA*speed_intercept
speed_slope ~~ NA*speed_slope

# set observed intercepts to zero - these should go into the latents
prod_t1 ~ 0
prod_t2 ~ 0
prod_t3 ~ 0
prod_t4 ~ 0
comp_t1 ~ 0 
log_rt_t1 ~ 0
log_rt_t2 ~ 0
log_rt_t3 ~ 0
log_rt_t4 ~ 0
log_rt_sd_t1 ~ 0 
log_rt_sd_t2 ~ 0
log_rt_sd_t3 ~ 0 
log_rt_sd_t4 ~ 0 
acc_t1 ~ 0 
acc_t2 ~ 0 
acc_t3 ~ 0 
acc_t4 ~ 0 
acc_sd_t1 ~ 0
acc_sd_t2 ~ 0
acc_sd_t3 ~ 0
acc_sd_t4 ~ 0
"

@@

```{r}
layout <- read.csv(here("misc/layout_dt3.csv"), header = FALSE)

graph_sem(model = fit3_long, text_size = 3, layout=t(layout))

@@

# Draft 4: Multi-level modeling of a single variable

Let's try NLME.

```{r}
d_sub_prod <- filter(d_sub, !is.na(prod)) |>
  group_by(subject_id) |>
  mutate(log_rt_0 = log_rt[1],
         acc_0 = long_window_accuracy[1])

library(nlme)

mod1 <- lme(prod ~ age, 
            random = ~ 1 | dataset_name / subject_id,
            na.action = na.omit, 
            data = d_sub_prod)

summary(mod1)
```

Plotting model.

```{r}
ages <- seq(min(d_sub_prod$age), max(d_sub_prod$age), .1)
datasets <- unique(d_sub_prod$dataset_name)

# fixed effects with CIs
newdata_global <- tibble(age = ages) 
global_preds <- emmeans(mod1, ~ age, at = newdata_global, CIs = TRUE) |> 
  summary() |>
  tibble()

# random effects
dataset_preds <- d_sub |>
  filter(!is.na(prod)) |> # filter on predictor
  group_by(dataset_name) |>
  summarise(min_age = min(age), 
            max_age = max(age)) |>
  group_by(dataset_name) |>
  mutate(df = map2(min_age, max_age, ~expand_grid(age = seq(.x, .y, .1)))) |>
  select(-min_age, -max_age) |>
  unnest(col = "df") |>
  ungroup()

dataset_preds$pred <- predict(mod1, 
                              newdata = dataset_preds, 
                              level = c(0,1))$predict.dataset_name

ggplot(d_sub_prod, aes(x = age, y = prod, col = dataset_name)) + 
  geom_point(alpha = .1) + 
  geom_line(data = global_preds, aes(y = emmean), col = "blue", size = 1.5) +
  geom_ribbon(data = global_preds, 
              aes(ymin = lower.CL, ymax = upper.CL, y = emmean), col = NA,
              fill = "blue",alpha = .1) + 
  geom_line(data = dataset_preds, aes(y = pred))
```

## Logistics

So I guess we need good parameters.

```{r}
d_sub$logistic = SSlogis(d_sub$age, Asym = 1, xmid = 0, scal = 3)

ggplot(d_sub, aes(x = age, y = prod)) + 
  geom_point(alpha = 1.) + 
  geom_line(aes(x = age, y = logistic))

Use a logistic growth model.

```{r}
mod2 <- nlme(model = prod ~ SSlogis(age, Asym = 1, xmid, scale),
             data = d_sub,
             fixed = xmid + scale ~ 1,
             random = xmid + scale ~ 1,
             groups = ~ dataset_name,
             start = c(xmid = 20, scale = 3),
             na.action = na.exclude, 
             control = list(maxIter = 1000, tolerance = 10))

summary(mod2)

# fixed effects (no CIs)
global_preds <- expand_grid(age = seq(min(d_sub_prod$age), 
                                      max(d_sub_prod$age), .1))
global_preds$pred <- predict(mod2, newdata = global_preds, level = 0)

# random effects
dataset_preds <- d_sub |>
  filter(!is.na(prod)) |> # filter on predictor
  group_by(dataset_name) |>
  summarise(min_age = min(age), 
            max_age = max(age)) |>
  group_by(dataset_name) |>
  mutate(df = map2(min_age, max_age, ~expand_grid(age = seq(.x, .y, .1)))) |>
  select(-min_age, -max_age) |>
  unnest(col = "df") |>
  ungroup()

dataset_preds$pred <- predict(mod2, 
                              newdata = dataset_preds, 
                              level = c(0,1))$predict.dataset_name

ggplot(d_sub_prod, aes(x = age, y = prod, col = dataset_name)) + 
  geom_point(alpha = .1) + 
  geom_line(data = global_preds, aes(y = pred), col = "blue", size = 1.5) +
  geom_line(data = dataset_preds, aes(y = pred))
```
## Interaction with t0 RT

Now let's use t0 processing attributes in a regression.

```{r}
d_sub_prod <- d_sub_prod |>
  group_by(subject_id) |>
  mutate(log_rt_0 = log_rt[1],
         acc_0 = long_window_accuracy[1])

mod3 <- nlme(model = prod ~ SSlogis(age, Asym = 1, xmid, scale),
             data = d_sub_prod,
             fixed = xmid + scale ~ log_rt_0,
             random = xmid + scale ~ 1,
             groups = ~ dataset_name,
             start = c(xmid = 20, scale = 3, `xmid:log_rt_0` = 0, `scale:log_rt_0` = 3),
             na.action = na.exclude, 
             control = list(maxIter = 1000, tolerance = 10))

summary(mod3)
```

```{r}
# fixed effects (no CIs)
global_preds <- expand_grid(age = seq(min(d_sub_prod$age), 
                                      max(d_sub_prod$age), .1), 
                            log_rt_0 = c(quantile(d_sub_prod$log_rt, 
                                                  c(.1,.25,.5,.75,.9), 
                                                  na.rm=TRUE)))
global_preds$pred <- predict(mod3, newdata = global_preds, level = 0)

# random effects
# dataset_preds <- d_sub |>
#   filter(!is.na(prod)) |> # filter on predictor
#   group_by(dataset_name) |>
#   summarise(min_age = min(age), 
#             max_age = max(age)) |>
#   group_by(dataset_name) |>
#   mutate(df = map2(min_age, max_age, ~expand_grid(age = seq(.x, .y, .1)))) |>
#   select(-min_age, -max_age) |>
#   unnest(col = "df") |>
#   ungroup()
#   
# dataset_preds$pred <- predict(mod3, 
#                               newdata = dataset_preds, 
#                               level = c(0,1))$predict.dataset_name

ggplot(d_sub_prod, aes(x = age, y = prod)) + 
  geom_point(alpha = .1) + 
  geom_line(data = global_preds, aes(y = pred, group = log_rt_0, col = as.factor(log_rt_0)))
# +
#   geom_line(data = dataset_preds, aes(y = pred))
```

## Interaction with accuracy

```{r}
mod4 <- nlme(model = prod ~ SSlogis(age, Asym = 1, xmid, scale),
             data = d_sub_prod,
             fixed = xmid + scale ~ log_rt_0 + acc_0,
             random = xmid + scale ~ 1,
             groups = ~ dataset_name,
             start = c(xmid = 20, scale = 3, `xmid:log_rt_0` = 0, `scale:log_rt_0` = 3, 
                       `xmid:acc_0` = 0, `scale:acc_0` = 3),
             na.action = na.exclude, 
             control = list(maxIter = 1000, tolerance = 10))

summary(mod4)
```
## power function with offset

So I guess we need good parameters.

```{r}
power_growth <- function(X, a, b, c)
{
  return((a * (X-c)^b))
}

d_sub_prod$power = power_growth(d_sub_prod$age, a = 1, b = 2.3, c = 12)

ggplot(d_sub_prod, aes(x = age, y = prod*680)) + 
  geom_point(alpha = 1.) + 
  geom_line(aes(x = age, y = power))

Now model.

```{r}
d_sub_prod$prod_680 <- d_sub_prod$prod * 680
mod5 <- nlme(model = prod_680 ~ power_growth(age, a, b, c),
             data = d_sub_prod,
             fixed = a + b + c ~ 1,
             random = a + b + c ~ 1,
             groups = ~ dataset_name,
             start = c(a = 1, b = 2.3, c = 12),
             na.action = na.exclude)

summary(mod5)

# fixed effects (no CIs)
global_preds <- expand_grid(age = seq(min(d_sub_prod$age), 
                                      max(d_sub_prod$age), .1))
global_preds$pred <- predict(mod5, newdata = global_preds, level = 0)

# random effects
dataset_preds <- d_sub |>
  filter(!is.na(prod)) |> # filter on predictor
  group_by(dataset_name) |>
  summarise(min_age = min(age), 
            max_age = max(age)) |>
  group_by(dataset_name) |>
  mutate(df = map2(min_age, max_age, ~expand_grid(age = seq(.x, .y, .1)))) |>
  select(-min_age, -max_age) |>
  unnest(col = "df") |>
  ungroup()

dataset_preds$pred <- predict(mod5, 
                              newdata = dataset_preds, 
                              level = c(0,1))$predict.dataset_name

ggplot(d_sub_prod, aes(x = age, y = prod_680, col = dataset_name)) + 
  geom_point(alpha = .1) + 
  geom_line(data = global_preds, aes(y = pred), col = "blue", size = 1.5) +
  geom_line(data = dataset_preds, aes(y = pred))
```

Having trouble fitting this model - throws an error sadly.

```{r eval=FALSE}
d_sub_prod$prod_681 <- d_sub_prod$prod_680 + 1
mod5 <- nlme(model = prod_681 ~ power_growth(age, a, b, c),
             data = d_sub_prod,
             fixed = a + b + c ~ log_rt_0,
             random = a + b + c ~ 1,
             groups = ~ dataset_name,
             start = c(a = 1, b = 2.3, c = 12, 
                       `a:log_rt_0` = 0.1, `b:log_rt_0` = 0.1, `c:log_rt_0` = 0.1),
             na.action = na.exclude)

summary(mod5)
```
